---
title: "Hospital Compare Website Assignment"
author: "Donato Scarano"
date: "21 giugno 2018"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction


The data for this assignment come from the Hospital Compare web site (http://hospitalcompare.hhs.gov)
run by the U.S. Department of Health and Human Services. The purpose of the web site is to provide data and
information about the quality of care at over 4,000 Medicare-certified hospitals in the U.S. This dataset essentially
covers all major U.S. hospitals. This dataset is used for a variety of purposes, including determining
whether hospitals should be fined for not providing high quality care to patients (see http://goo.gl/jAXFX
for some background on this particular topic).
The Hospital Compare web site contains a lot of data and we will only look at a small subset for this
assignment. The zip file for this assignment contains three files
• outcome-of-care-measures.csv: Contains information about 30-day mortality and readmission rates
for heart attacks, heart failure, and pneumonia for over 4,000 hospitals.
• hospital-data.csv: Contains information about each hospital.
• Hospital_Revised_Flatfiles.pdf: Descriptions of the variables in each file (i.e the code book).
A description of the variables in each of the files is in the included PDF file named Hospital_Revised_Flatfiles.pdf.

##Execution

Read the outcome data into R via the read.csv function and look at the first few rows.

1. Set working directory

```{r}
setwd("C:/Users/dscarano/Downloads")
```
2. Read the outcome data

```{r}
outcome <- read.csv("outcome-of-care-measures.csv", colClasses = "character")
head(outcome)
```

There are many columns in this dataset.
you can see the number of rows/columns with the nrow/ncol function:

```{r}
ncol(outcome) 
```

In addition, you can see the names of each column by typing:

```{r} 
names(outcome)
```

To make a simple histogram of the 30-day death rates from heart attack (column 11 in the outcome dataset)

```{r}
outcome[, 11] <- as.numeric(outcome[, 11])
hist(outcome[, 11])
```

#Finding the Best Hospital in a state

```{r}
best <- function(state, outcome) {
        ## Read outcome data
        data <- read.csv("outcome-of-care-measures.csv")
        
        ## Check that state and outcome are valid
        states <- levels(data[, 7])[data[, 7]]
        state_flag <- FALSE
        for (i in 1:length(states)) {
                if (state == states[i]) {
                        state_flag <- TRUE
                }
        }
        if (!state_flag) {
                stop ("invalid state")
        } 
        if (!((outcome == "heart attack") | (outcome == "heart failure")
            | (outcome == "pneumonia"))) {
                stop ("invalid outcome")
        }
        
        ## Return hospital name in that state with lowest 30-day death rate
        col <- if (outcome == "heart attack") {
                11
        } else if (outcome == "heart failure") {
                17
        } else {
                23
        }
        
        data[, col] <- suppressWarnings(as.numeric(levels(data[, col])[data[, col]]))
        data[, 2] <- as.character(data[, 2])
        statedata <- data[grep(state, data$State), ]
        orderdata <- statedata[order(statedata[, col], statedata[, 2], na.last = NA), ]
        orderdata[1, 2]
}

```

# Ranking Hospitals by outcome in a state

```{r}
rankhospital <- function(state, outcome, num = "best") {
        ## Read outcome data
        data <- read.csv("outcome-of-care-measures.csv")

        ## Check that state and outcome are valid
        states <- levels(data[, 7])[data[, 7]]
        state_flag <- FALSE
        for (i in 1:length(states)) {
                if (state == states[i]) {
                        state_flag <- TRUE
                }
        }
        if (!state_flag) {
                stop ("invalid state")
        } 
        if (!((outcome == "heart attack") | (outcome == "heart failure")
              | (outcome == "pneumonia"))) {
                stop ("invalid outcome")
        }

        ## Return hospital name in that state with the given rank 30-day death 
        ## rate
        col <- if (outcome == "heart attack") {
                11
        } else if (outcome == "heart failure") {
                17
        } else {
                23
        }
        
        data[, col] <- suppressWarnings(as.numeric(levels(data[, col])[data[, col]]))
        data[, 2] <- as.character(data[, 2])
        statedata <- data[grep(state, data$State), ]
        orderdata <- statedata[order(statedata[, col], statedata[, 2], na.last = NA), ]
        if(num == "best") {
                orderdata[1, 2]
        } else if(num == "worst") {
                orderdata[nrow(orderdata), 2]
        } else{
                orderdata[num, 2]
        }
}
```
#Ranking Hospital in all States


```{r}
rankall <- function(outcome, num = "best") {
        ## Read outcome data
        data <- read.csv("outcome-of-care-measures.csv")
        
        ## Check that outcome is valid
        if (!((outcome == "heart attack") | (outcome == "heart failure")
              | (outcome == "pneumonia"))) {
                stop ("invalid outcome")
        }

        ## For each state, find the hospital of the given rank
        col <- if (outcome == "heart attack") {
                11
        } else if (outcome == "heart failure") {
                17
        } else {
                23
        }
        
        data[, col] <- suppressWarnings(as.numeric(levels(data[, col])[data[, col]]))
        data[, 2] <- as.character(data[, 2])
        
        # Generate an empty vector that will be filled later, row by row, to 
        # generate the final output.
        output <- vector()
        
        states <- levels(data[, 7])
        
        for(i in 1:length(states)) {
                statedata <- data[grep(states[i], data$State), ]
                orderdata <- statedata[order(statedata[, col], statedata[, 2], 
                                             na.last = NA), ]
                hospital <- if(num == "best") {
                        orderdata[1, 2]
                } else if(num == "worst") {
                        orderdata[nrow(orderdata), 2]
                } else{
                        orderdata[num, 2]
                }
                output <- append(output, c(hospital, states[i]))
        }

        ## Return a data frame with the hospital names and the (abbreviated) 
        ## state name
        output <- as.data.frame(matrix(output, length(states), 2, byrow = TRUE))
        colnames(output) <- c("hospital", "state")
        rownames(output) <- states
        
        output
}
